PHAT RAID, yo

### 2014-05-15 paper/paper.tex
future work
Update the paper

Our setting is in a data center where we have low network latency,
high network throughput, and the need to store large files with tolerance to
some CPU, network, and disk failure. We reason that with these characteristics,
disk read/write latency will be the major factor in overall latency to store a file
with replication. As a result, we borrow ideas from RAID that allow for faster
disk access, and mix them with ideas for distributed consensus algorithms that
allow for consistent replication.

@@

\phatraid{} is implemented in Erlang. Erlang is a pure, functional language with
extensive library support for distributed applications. Erlang has one unit of
concurrency, the \emph{process}, and one unit of distribution, the
\emph{node}. Each Erlang node consists of an arbitrary number of
processes. Erlang nodes may live on the same machine or on different machines
connected by a network. Every Erlang node is given a name by an Erlang name
server. Erlang processes may communicate with any Erlang node or process it can
name. Erlang processes interact with one another by making RPC calls or sending
messages. Identifiers beginning with a lowercase letter are actually atoms, like
Scheme's symbols. Identifiers beginning with an uppercase letter are normal
identifiers.

@@

On a client node, the raid-client spawns $C$ client processes which each
maintain a connection to one of the $C$ \phat{} clusters. Each client process
mediates all communication between its assigned \phat{} cluster and the
raid-client.

@@

@@

The server cluster is a collection of nodes running the same server code. One of
these nodes is designated the master node (aka \phat{} Master). The other nodes are
known as \phat{} replicas or just replicas.

The master and each replica has four components: the supervisor, the
(unfortunately named) server, VR, and the file system:

@@

Additionally, the \texttt{gen\_server} and \texttt{gen\_fsm} behaviors allow
the programmer to specify an arbitrary state value which will be passed around
\`{a} la functional reactive programming.

@@

Listing \ref{lst:supervisor} is taken from our supervisor code. When a
supervisor is started, the \texttt{supervisor} behavior looks for a unary
procedure called \lstinline!init!. The \lstinline!init! procedure returns a
child specification which is a pair of a restart specification and a list of
children. In Listing \ref{lst:supervisor}, the restart specification,
\lstinline!{one_for_all,1,5}! specifies that if any one node dies, all nodes
should be restarted, unless more than one failure has occurred in the past five
seconds. If more than one failure occurs in five seconds, the supervisor kills
all the children and shuts itself down as well.

@@

A state in the \texttt{gen\_fsm} behavior manifests as a procedure which accepts
two arguments: the incoming message and the store\footnote{Mutable state is
  implemented as a pure store value which is passed around by the behavior}. The
return value of a state-procedure is usually the 4-tuple
\lstinline!{next_state,NextState,NewStore,Timeout}!. The \texttt{gen\_fsm}
behavior will transition to \lstinline!NextState! and set the store to
\lstinline!NewStore!. When a new message is received the \texttt{gen\_fsm}
behavior will invoke the \lstinline!NextState! procedure with the new message
and the \lstinline!NewStore!.

The \phat{} file system achieves consensus via the Viewstamped Replication
protocol \cite{liskov2012viewstamped}. Listing \ref{lst:genfsm} defines how to
handle prepare messages sent to a node in the replica state. It is piecewise
defined by pattern matching on the arguments. The first two definition clauses
use the \lstinline!when! keyword to specify arbitrary required relationships
between the matched variables.

@@

Listing \ref{lst:processing2} handles messages from the current master. The
\lstinline!prepareBuffer! stores prepare messages that arrived out of order. In
particular, if the log ends at operation number $n$ and message $n+1$ is dropped
by the network, all subsequent messages will be buffered. The buffered messages
will not be added to the log until message $n+1$ is received. The call to
\lstinline!processBuffer! moves messages to the log, if all previous messages
have now been received. The call to \lstinline!processLog! commits logged
messages which now have been committed by the master node. Lines 405 to 408
destructure \lstinline!AfterLog! and bind variables for later use. The final
\lstinline!if! statement responds to the master unless the master has committed
messages we haven't received.\footnote{This can happen if the network drops a
  message whose operation number lies between the replica's last commit number
  and the master's last commit number.}

@@

Four arrangements of failed nodes in \phatraidcf{3}{1} are depicted in Figure
\ref{fig:failures}. In this section, if we claim ``data is lost'', we mean that
if no persistent store is used, data will be lost. Adding a persistent store
obviously prevents loss of committed data. Additionally, we assume node failures
are independent and uniformly distributed throughout the \phatraid{} group.

@@

The top right corner, Figure \ref{fig:failures_noprogress}, depicts four of the
nine nodes failing. This state is a failure state. The group cannot make
progress; however, no data can been lost. Progress is blocked because more than
one cluster cannot reach consensus. Since no cluster was completely destroyed,
all data still exists on at least one node.

The bottom left corner, Figure \ref{fig:failures_impossible}, is impossible
because if data can be lost then we certainly cannot make progress.

@@

In general, progress ceases if more than one cluster \emph{cannot reach
  consensus} and data loss occurs if more than one \emph{cluster completely
  fails}. The lower bound of failed nodes for progress cessation in
\phatraidcf{$C$}{$f$} is:

@@

By the pidgeon hole principle, the upper bound of failed nodes which still
permit progress is,

    $$ C\cdot f + 1 $$

because we can place $f$ failures in all but one cluster and place $f+1$
failures in the final cluster.

The upper bound for no data loss is,

    $$ C\cdot 2f + 1 $$

because we can keep one node of all but one cluster alive. From these $C-1$
nodes we can reconstruct the entire file system.

The bounds presented above cannot fully describe the failure threshold of a
\phatraid{} cluster because of the interplay between VR and RAID. Instead, we
will discuss the probability of group failure given $n$ node failures.

@@

When five nodes fail, they can be distributed into each group: (3,1,1), (3,2,0)
and (2,2,1). The first case does not stop progress because only one cluster
fails. The later two cases stop progress because two clusters fail. We calculate
the probability of \phatraid{} group failure when five nodes fail:

@@

The general formula for the failure probability distribution is not clear, but
it is clear that a \phatraidcf{$C$}{$f$} group is absolutely resilient to $2(f +
1) - 1$ or fewer failures. A \phatraidcf{$C$}{$f$} group is likely resilient
to $2(f+1)$ failures and unlikely resistant to $C\cdot f + 1$ failures.

For exactly $2(f+1)$ node failures, the probability of \phatraidcf{$C$}{$f$}
group failure is:

@@

\begin{figure}
  \includegraphics[scale=0.45]{Store.pdf}
  \caption{Scatter Plot of Various \phatraidcf{$C$,$f$} Configurations}
  \label{fig:plot}
\end{figure}

We evaluated a prototype \phatraidcf{C}{f} implementation. We simulated disk
latency with sleeps. All experiments are run on a 2013-model MacBook Pro with
8GB of RAM, and a 3GHz Intel Core i7. We found that latency scales as
$\frac{1}{C-1}$. This is the expected behavior for breaking a file into $C-1$
chunks. We always use one bit of parity.

@@

We would like to perform a real evaluation rather than a simulation. 
It seems that as file size increases, PhatRaid becomes more appealing 
for lower commit latency. It is not clear from our
simulation, however, when the file size is small enough for other effects (such as
network latency, communications overhead) to become more important. And, as a 
result, PhatRaid having no improvements (or possibly worse performance).

>>>>

Our setting is in a data center where we have low network latency, high network
throughput, and the need to store large files with tolerance to some CPU,
network, and disk failure. In this setting, we reason that the major bottleneck
for replicated file storage is disk read/write latency. We combat this
bottleneck with ideas from RAID, which allow for faster disk access, and ideas
from distributed consensus algorithms, which allow for consistent replication.

@@

\phatraid{} is implemented in Erlang. Erlang is a pure, functional language with
extensive library support for distributed applications. Erlang has one unit of
concurrency, the \emph{process}, and one unit of distribution, the
\emph{node}. Each Erlang node consists of an arbitrary number of
processes. Erlang nodes may live on the same machine or on different machines
connected by a network. Every Erlang node is given a name by an Erlang name
server. Erlang processes may communicate with any Erlang node or process it can
name. Erlang processes interact with one another by making RPC calls or sending
messages. Identifiers beginning with a lowercase letter are actually atoms, like
Scheme and Lisp's symbols. Identifiers beginning with an uppercase letter are
normal identifiers.

@@

On a client node, the raid-client spawns $C$ client processes which each
maintain a connection to one of the $C$ \phat{} clusters. Each client process
mediates communication between its assigned \phat{} cluster and the raid-client.

@@

\subsection{\phat{} Clusters}

A server cluster is a collection of nodes which maintain a distributed file
system. Each server cluster designates a unique master node (aka the \phat{}
Master). The other nodes are known as \phat{} replicas or just replicas.

The master and each replica have four components: the supervisor, the server,
VR, and the file system:

@@

@@

Additionally, the \texttt{gen\_server} and \texttt{gen\_fsm} behaviors allow the
programmer to specify an arbitrary state value which will be passed around \`{a}
la functional reactive programming. We call this state value the \emph{store}.

@@

Listing \ref{lst:supervisor} is taken from our supervisor code. When a
supervisor is started, the \texttt{supervisor} behavior looks for a procedure
called \lstinline!init!. The \lstinline!init! procedure returns a child
specification, which is a pair of a restart specification and a list of
children. In Listing \ref{lst:supervisor}, the restart specification,
\lstinline!{one_for_all,1,5}! specifies that if any one node dies, all nodes
should be restarted, unless more than one failure has occurred in the past five
seconds. If more than one failure occurs in five seconds, the supervisor kills
all its children and then shuts itself down.

@@

A state in the \texttt{gen\_fsm} behavior manifests as a procedure which accepts
two arguments: the incoming message and the store. The return value of a
state-procedure is usually the 4-tuple
\lstinline!{next_state,NextState,NewStore,Timeout}!. The \texttt{gen\_fsm}
behavior will transition to \lstinline!NextState! and set the store to
\lstinline!NewStore!. When a new message is received the \texttt{gen\_fsm}
behavior will invoke the \lstinline!NextState! procedure with the new message
and the \lstinline!NewStore!.

The \phat{} file system reaches consensus via the Viewstamped Replication
protocol \cite{liskov2012viewstamped}. Listing \ref{lst:genfsm} defines how to
handle prepare messages sent to a node in the replica state. It is piecewise
defined by pattern matching on the arguments. The first two definition clauses
use the \lstinline!when! keyword to specify arbitrary required relationships
between the matched variables.

@@

Listing \ref{lst:processing2} handles messages from the current master. The
\lstinline!prepareBuffer! stores prepare messages that arrived out of order. In
particular, if the log ends at operation number $n-1$ and message $n$ is dropped
by the network, all subsequent messages, $n+i$ will be buffered. The buffered
messages will not be added to the log until message $n$ is received.

The call to \lstinline!processBuffer! moves messages to the log, if all previous
messages have now been received. The call to \lstinline!processLog! commits
logged messages which have been newly committed by the master node. Lines
405--408 destructure \lstinline!AfterLog! and bind variables for later use. The
final \lstinline!if! statement responds to the master unless the master has
committed messages we haven't yet received.\footnote{This can happen if the
  network drops a message whose operation number lies between the replica's last
  commit number and the master's last commit number.}

@@

Four arrangements of failed nodes in \phatraidcf{3}{1} are depicted in Figure
\ref{fig:failures}. In this section, ``data is lost'' means that, if no
persistent, fail-safe store is used, some data will be lost. Adding a
persistent, fail-safe store obviously prevents loss of committed
data. Additionally, we assume node failures are independent and uniformly
distributed throughout the \phatraid{} group.

@@

The top right corner, Figure \ref{fig:failures_noprogress}, depicts four of the
nine nodes failing. This state is a failure state. The group cannot make
progress; however, no data can been lost\footnote{Here and later in this paper
  we mean that no data which has been decided and propagated throughout the
  cluster can be lost. In VR, by contrast, if data is decided and propagated
  throughout the cluster, that data can only be lost if every node is lost. We
  also assume the nodes do not have persistent, fail-safe stores.}. Progress is
blocked because more than one cluster cannot reach consensus. Since no cluster
was completely destroyed, all data still exists on at least one node.

The bottom left corner, Figure \ref{fig:failures_impossible}, is impossible
because the group certainly cannot make progress if data has been lost.

@@

In general, progress ceases if more than one cluster \emph{cannot reach
  consensus} and data loss occurs if more than one cluster \emph{completely
  fails}. The lower bound of failed nodes for progress cessation in
\phatraidcf{$C$}{$f$} is:

@@

By the pidgeon hole principle, the upper bound of failed nodes which still
permit progress is:

    $$ (C-1)\cdot f + 2f + 1 $$

because $C-1$ clusters can make progress, i.e. at least $f+1$ nodes are up in
each cluster, and only one cluster is completely dead.

The upper bound for no data loss is:

    $$ (C-1)\cdot 2f + 2f + 1 $$

because $C-1$ clusters have at least one node up and one cluster has no nodes
up.

The bounds presented above cannot fully describe the failure threshold of a
\phatraid{} cluster because of the interplay between VR and RAID. Instead, we
discuss the probability of group failure given $n$ node failures.

@@

When five nodes fail, they can be distributed into each group in three different
ways: (3,1,1), (3,2,0) and (2,2,1). The first case does not stop progress
because only one cluster fails. The later two cases stop progress because two
clusters fail. We calculate the probability of \phatraid{} group failure when
five nodes fail:

@@

The general formula for the failure probability distribution is not clear, but
it is clear that a \phatraidcf{$C$}{$f$} group is absolutely resilient to $2(f +
1) - 1$ or fewer failures. A \phatraidcf{$C$}{$f$} group is likely resilient
to $2(f+1)$ failures and unlikely resistant to $(C-1)\cdot f + 2f + 1$ failures.

For exactly $2(f+1)$ node failures, the probability of a \phatraidcf{$C$}{$f$}
group failure is:

@@

\begin{figure}
  \includegraphics[scale=0.45]{Store.pdf}
  \caption{Scatter Plot of Various \phatraidcf{$C$}{$f$} Configurations}
  \label{fig:plot}
\end{figure}

We evaluated a prototype \phatraidcf{$C$}{$f$} implementation. We simulated disk
latency with sleeps. All experiments are run on a 2013-model MacBook Pro with
8GB of RAM, and a 3GHz Intel Core i7. We found that latency scales as
$\frac{1}{C-1}$. This is the expected behavior for breaking a file into $C-1$
chunks. We always use one bit of parity.

Figure \ref{fig:plot} is a scatter plot of the time to store a 100 MB file in a
\phatraid{} group. Each point represents a different configuration. The x-axis
indicates the number of nodes needed for that configuration. The x-axis can also
be seen as a measure of cost to create the configuration.

The slow linear growth from \phatraidcf{1}{1} to \phatraidcf{1}{14} reveals the
cost of communication in large \phat{} clusters. This phenomenon is also present
in the slight increase in cost from \phatraidcf{$C$}{1} to \phatraidcf{$C$}{2}.

@@

We would like to perform a real evaluation rather than a simulation. 
It seems that as file size increases, PhatRaid becomes more appealing 
for lower commit latency. It is not clear from our
simulation, however, when the file size is small enough for other effects (such as
network latency, communications overhead) to become more important.
